A User-Extensible and Adaptable Parser Architecture

### 2008-12-08 doc/sgai-2008-slides.tex
Improvements after dry runs with Carl.

\institute[Trinity College]
{
    School of Computer Science and Statistics,\\
    Trinity College, University of Dublin
}

@@

\begin{frame}
    \timingnote{30 seconds}
    \titlepage{}
\end{frame}

\begin{frame}{Outline Of This Talk}
    \timingnote{30 seconds --- 1 minute}
    \tableofcontents{}
\end{frame}

@@

        \item [However] Rejecting mail at SMTP-time is more efficient, and
            senders of non-spam mails that are incorrectly rejected will
            receive a bounce notification immediately.

        \item [Project Aim] To improve SMTP-time anti-spam techniques, by
            providing a platform for reasoning about anti-spam techniques.

        \item [Method] Parse Postfix log files and analyse the
            extracted data.

@@

        \item [Framework] The framework manages rules and provides support
            functions to actions, e.g.\ data storage, inter-action
            communication, and utility functions.  For each input log line,
            the framework tries each rule in turn until one matches, then
            it invokes the action specified by the rule.

        \item [Actions] Each action performs the processing required by one
            type of log line.  Actions correspond to events: events within
            Postfix create log lines, and actions perform the processing
            required by those events.

        \item [Rules] Each rule matches one log line variant, e.g.\ the log
            line resulting from client connecting, a mail successfully
            delivered, or a mail rejected by an anti-spam measure.

@@

        \item Rules can be tagged with conditions which must hold before
            the rule will be used when attempting to match an input.

        \item The parser does not need to be recompiled when rules are
            changed, allowing: quicker testing of modified rules; easing
            interchange of rulesets; combining two or more rulesets.

@@

        \item Rules can be added, changed, or removed by actions while the
            parser is running, allowing dynamically changing parsing, e.g.\
            macros.  This feature must be used with care.
            \timingnote{Currying functions in a calculator might be a good
            example.}

@@

        \item 169 rules parsing 522 contiguous log files: the top 10 rules
            match 85.03\% of the log lines, the remainder tail off similar
            to a Power Law distribution.

        \item 18 actions; most are triggered by only one or two rules.
            \timingnote{client connecting, postfix reloading}

        \item The top three actions are triggered by 41, 49, and 59 rules.
            \timingnote{build/stats-action-distribution:}
            \timingnote{uninteresting, save\_data, delivery\_rejected}

@@

        \item Median throughput is 80.85~MB (480,569 log lines) parsed per
            minute; median log file size is 50 MB, containing 285,000 log
            lines.

@@

        \item Rules are sorted by how frequently they match, resulting in a
            mean reduction of 14.78\% in parsing time compared to using
            unsorted rules.  Priority attribute can be used to override
            this.

        \item The architecture scales extremely well as the number of rules
            increases: a 46.05\% increase in the Postfix ruleset ($115
            \rightarrow{} 169$ rules) resulted in a mean increase of 0.63\%
            in parsing time.

        \item Parsing time scales linearly with input size, but
            \textit{is\/} affected by input content: the content determines
            the actions that will run, and some actions take longer to run
            than others.

@@

\section{Future Directions}

\begin{frame}{Future Directions}

@@

@@

        \item Use Machine Learning techniques instead of writing rules;
            Instance Based Learning techniques must be used due to the size
            of the data set.  \timingnote{Compare to SLCT-based tool.}

        \item Detection of overlapping rules: intersection of Finite
            Automata or Push-Down Automata.

@@

        \item Parser Architecture makes it easy to develop a flexible yet
            strict parser.

        \item Parser Architecture makes it easy for users to parse new or
            changed inputs.

        \item Parser can still be efficient and fast.

>>>>

\institute[TCD]
{
    School of Computer Science and Statistics,\\
    Trinity College, University of Dublin
}

@@

\begin{frame}
    \timingnote{30 seconds}
    \titlepage{}
    \timingnote{Mention paper in journal}
\end{frame}

\begin{frame}{Outline Of This Talk}
    \timingnote{30 seconds --- 1 minute\newline{}}
    \timingnote{Brief statement of motive during background}
    \tableofcontents{}
\end{frame}

@@

        \item [However] Rejecting mail at SMTP-time (during the delivery
            from sender to recipient) is more efficient, and senders of
            non-spam mails that are incorrectly rejected will receive a
            bounce notification immediately.

        \item [Project Aim] To improve SMTP-time anti-spam techniques.

        \item [Method] Parse Postfix log files and analyse the
            extracted data.  \timingnote{Explain where log lines come from}

@@

        \item [Framework] The framework manages rules, manages the parsing
            process, and provides support functions to actions.  For each
            input log line, the framework tries each rule in turn until one
            matches, then it invokes the action specified by the matching
            rule.

        \item [Actions] Each action performs the processing required by one
            \alert{type} of log line.  Actions correspond to events: events
            within Postfix create log lines, and actions perform the
            processing required by those events.

        \item [Rules] Each rule matches one log line \alert{variant}; a
            single log line \alert{type} may have many log line
            \alert{variants}, e.g.\ rejection of a mail delivery attempt
            has 59 variants.

@@

        \item Rules can be tagged with conditions which must hold before
            the rule will be used when attempting to match an input.
            \timingnote{Simple, but could be extended as necessary.}

        \item Parsers and rulesets remain separate.

@@

        \item Rules can be added, changed, or removed by actions while the
            parser is running, allowing dynamically changing parsing, e.g.\
            macros.  \timingnote{Currying functions in a calculator might
            be a good example.}

@@

        \item 115 (now 169) rules parsing 93 (now 522) contiguous log
            files: the 10 most frequently matching rules match 85.03\% of
            the log lines, the remainder tail off similar to a Power Law
            distribution.

        \item 18 actions; 15 are triggered by only one or two rules.
            \timingnote{client connecting, postfix reloading}

        \item 3 actions are triggered by an average of 50 rules:
            uninteresting (41), save\_data (49), and delivery\_rejected
            (59).
            %\timingnote{build/stats-action-distribution:}

@@

        \item Median throughput is 80.85~MB (480,569 log lines) parsed per
            minute; median log file size is 50 MB, containing 285,000 log
            lines; typically 6 or 7 log lines per mail.

@@

        \item Rules are sorted by how frequently they match, resulting in a
            mean reduction of 14.78\% in parsing time.  Priority attribute
            can be used to override this.  \timingnote{change in ordering
            is informative.  Described in paper}

        \item The architecture scales extremely well as the number of rules
            increases: a 46.05\% increase in the Postfix ruleset ($115
            \rightarrow{} 169$ rules) resulted in a mean increase of 0.63\%
            in parsing time.  \timingnote{less than 2/3 of 1\%; 2 orders of
            magnitude}

        \item Parsing time scales linearly with input size, but
            \textit{is\/} affected by input log lines: the log lines
            determine the number of rules used and the actions that will
            run, and some actions take longer to run than others.

@@

\section{Ongoing And Future Directions}

\begin{frame}{Ongoing And Future Directions}

@@

    Ongoing:

@@

    Future:

        \item Use Machine Learning techniques instead of writing rules;
            Instance Based Learning techniques must be used due to the size
            of the data set.  This is one use of the ability to add or
            change rules at run-time.  \timingnote{Compare to SLCT-based
            tool.}

        \item Detection of overlapping rules: intersection of Finite
            Automata or Push-Down Automata; how do rule conditions overlap?

@@

        \item This Parser Architecture makes it easy to develop a flexible
            yet strict parser.

        \item This Parser Architecture nevertheless makes it easy for users
            to parse new or changed inputs.

        \item Parsers written using this architecture can still be
            efficient and fast.

        \item The Postfix log file parser provides a basis for evaluating
            SMTP-time anti-spam techniques, something that has not been
            available before.
